---
editor: 
  markdown: 
    wrap: 72
---

# Data

## Description

### Data Source

We obtained the data from the United States [Census Bureau
Website](https://www.census.gov/programs-surveys/nsch/data/datasets.html),
which is an official website of the US government. The data is collected
from 2022 NSCH (National Survey of Children's Health), a national and
state-level survey on information related to the physical and mental
health of children from 0 to 17 years old. The survey is conducted by
the United States Census Bureau, Associate Director for Demographic
Programs on behalf of the United States Department of Health and Human
Services (HHS), Health Resources and Services Administration's (HRSA)
Maternal and Child Health Bureau (MCHB) .

### Format of the data, Frequency of updates, Dimensions

The 2022 Topical Data is originally in SAS format. The NSCH is conducted
every year, so the data is collected and released every year.
Specifically, the NSCH 2022 was collected from July 2022 to January
2023. The data has two dimensions. Each row represents a sample, and
each column represents the answers to each question in the
questionnaires. The 2022 Topical Data file includes records for 54,103
children (54103 rows), and there are answers to 490 questions for each
child (490 columns). Issues / problems with the data The data is
collected from questionnaires, so some information may not be complete.
The 2022 topical data file involves both the complete and sufficient
partial topical questionnaires. Therefore, we need to deal with missing
values before we conduct analyses. Besides, the data involves different
aspects of children's health and well-being, but our purpose is to focus
on children's mental health. So we need to carefully identify and choose
the relevant information from the 490 columns.\

### Import the data

The original data is in SAS format, and it is not that easy to analyze.
Thus, we plan to convert the data to csv format manually and store it as
"nsch.csv". Then, we just need to import the "nsch.csv" file into
R-studio in DataFrame format to prepare the data for analyses.

## Research Plan

### Census Question Columns

Based on the **4 big categories of research questions**, we select
different census questions to explore:

-   **Parents / Environment**

    `A1_ACTIVE, A1_AGE, A1_BORN, A1_EMPLOYED, A1_GRADE, A1_MARITAL, A1_MENTHEALTH, A1_PHYSHEALTH, FAMILY_R, HHCOUNT, HIGRADE_TVIS, HOMEEVIC, HOPEFUL`

    These variables are basic information of guardians like whether
    they've served on active duty, the education level, life attitude,
    physical and mental health self-evaluation. Living environments like
    family structure and household sizes can be also considered when
    evaluating the correlation with mental health issues.

-   **Children's Experience**

    `ACE1, ACE10, ACE11, ACE12, ACE3, ACE4, ACE5, ACE6, ACE7, ACE8, ACE9, BULLIED_R, BULLY, EVERHOMELESS, GRADES`

    In this category, guardians are asked to unveil their child's
    traumatic experience like being treated unfairly and living in an
    unsettled environment. In addition, experience of peer relationship
    (bully, grade) are also subjects to be tested.

-   **Physical Health**

    `FASD, GENETIC, HOURSLEEP, PHYSACTIV, PHYSICALPAIN`

    By accessing the relation between mental health issues and physical
    health factors like activity frequency and hours of sleep, we can
    answer the last category of questions.

-   **Mental Health Indicator**

    `K2Q30A, K2Q31A, K2Q32A, K2Q33A, K2Q34A, K2Q35A, K2Q37A, K2Q38A`

    These are the target variables, which includes YES/NO questions on
    whether this child is diagnosed to have mental health issues.

-   **Treatment Indicator**

    `K4Q22_R, K4Q23`

    Information of whether this child has taken treatments/medication or
    not.

### Visualizations

Since almost all questions in this census have categorical answers, we
can use contingency tables to draw stacked bar charts, Cleveland dot
plots and mosaic plots to see the correlation. As shown above, we kept a
large set of variables to explore, and we might want to try PCA as well
to reduce dimension.

*Please refer to the [official variable
list](https://www2.census.gov/programs-surveys/nsch/technical-documentation/codebook/2022-NSCH-Topical-Variable-List.pdf)
provided on the website for detailed question prompts.*

## Missing value analysis

```{r}
nsch_df <- read.csv("data/nsch.csv")
missing <- is.na(nsch_df)
dim(nsch_df)
dim(nsch_df[apply(missing, 1, any), ])
```

The dataset is large enough even if we choose removing all the NAs.
However, we still want to figure out the pattern of missing values to
avoid misinterpreation of data. We first check the number of
columns/rows containing missing values using `colSums()` and
`rowSums()`.

### `colSums()`

```{r}
col_sum_na <- colSums(missing) |> sort(decreasing = TRUE)
col_sum_na
barplot(col_sum_na, cex.names=0.7, las=2, main="Missing Value Counts of Each Column")
```

We can see that columns starts with "ACE" tend to have more missing
values. If we check the meaning of variables, we can see that these
columns are questions related to "Child Experienced". The question
prompt starts with "To the best of your knowledge", and only 2 options
(1=Yes or 2=No) are provided. People who take the questionaire might
take this question as an optional one and leave it blank when they are
unsure about the answer. If our inference here is correct, we can drop
those rows without concerning serious effects.

### Frequency of `rowSums()`

For other columns, we would want to look at the `rowSums()` frequency.

We are going to first remove columns starts with "ACE" (see previous
section for analysis) and columns with 0 missing values. Then we filter
to only include rows containing `NA`. We can see the dimension of
missing value dataframe:

```{r}
nsch_df_except_ace <- nsch_df[,-seq(16,25)]
col_wo_na <- c("HHID", "A1_GRADE", "HHCOUNT", "HIGRADE_TVIS")
col_names <- names(nsch_df_except_ace)
nsch_df_except_ace <- nsch_df_except_ace[,-which(col_names %in% col_wo_na)]
missing_except_ace <- is.na(nsch_df_except_ace)
nsch_df_except_ace <- nsch_df_except_ace[apply(missing_except_ace, 1, any), ]
dim(nsch_df_except_ace)
```

Then, we can draw the frequency bar chart of the missing vale of each
row.

```{r}
freq_table_na_each_row <- rowSums(is.na(nsch_df_except_ace)) |> table()
barplot(freq_table_na_each_row, cex.names=0.6,
        main="Frequency of Missing Value Counts of Each Row")
```

The values represented by each bin are:

```{r}
freq_table_na_each_row
```

After ignoring missing value in "ACE" columns, we can see among 3547
rows containing missing values, 2636 of them only have 1 value missing.
This indicates that it is unlikely to have causal relationship between
missing values in columns. We are going to validate this statement by
drawing a correlation matrix on missing values.

### Correlation Matrix

Using an indicator dataframe, we put 1 on the position where the
truncated dataframe has `NA` in that position, and put 0 otherwise. We
are going to draw a correlation matrix of this indicator to see whether
missing a value in column A lead to missing a value in column B.

```{r}
cor_matrix <- cor(as.data.frame(missing_except_ace) +1-1)

library(ggplot2)
# Reshape the correlation matrix for ggplot
cor_df <- as.data.frame(as.table(cor_matrix))
names(cor_df) <- c("Var1", "Var2", "Corr")

# Plot the correlation matrix using ggplot2
heatmap <- ggplot(cor_df, aes(Var1, Var2, fill = Corr)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_gradient2(low = "blue", high = "red", mid="white",
                       name = "Correlation", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Matrix")

# Display the plot
print(heatmap)
```

We can see no extreme correlation is found. Given that the dataset is
large enough and most columns are categorical, we choose to remove
missing values.
