# Data

## Description

### Data Source

We obtained the data from the United States [Census Bureau
Website](https://www.census.gov/programs-surveys/nsch/data/datasets.html),
which is an official website of the US government. The data is collected
from 2022 NSCH (National Survey of Children's Health), a national and
state-level survey on information related to the physical and mental
health of children from 0 to 17 years old. The survey is conducted by
the United States Census Bureau, Associate Director for Demographic
Programs on behalf of the United States Department of Health and Human
Services (HHS), Health Resources and Services Administration's (HRSA)
Maternal and Child Health Bureau (MCHB) .

### Format of the data, Frequency of updates, Dimensions

The 2022 Topical Data is originally in SAS format. The NSCH is conducted
every year, so the data is collected and released every year.
Specifically, the NSCH 2022 was collected from July 2022 to January
2023. The data has two dimensions. Each row represents a sample, and
each column represents the answers to each question in the
questionnaires. The 2022 Topical Data file includes records for 54,103
children (54103 rows), and there are answers to 490 questions for each
child (490 columns). Issues / problems with the data The data is
collected from questionnaires, so some information may not be complete.
The 2022 topical data file involves both the complete and sufficient
partial topical questionnaires. Therefore, we need to deal with missing
values before we conduct analyses. Besides, the data involves different
aspects of children's health and well-being, but our purpose is to focus
on children's mental health. So we need to carefully identify and choose
the relevant information from the 490 columns.\

### Import the data

The original data is in SAS format, and it is not that easy to analyze.
Thus, we plan to convert the data to csv format manually and store it as
"nsch.csv". Then, we just need to import the "nsch.csv" file into
R-studio in DataFrame format to prepare the data for analyses.


## Missing value analysis

```{r}
nsch_df <- read.csv("data/nsch.csv")
missing_df <- nsch_df[apply(is.na(nsch_df), 1, any), ]
dim(nsch_df)
dim(missing_df)
```

The dataset is large enough even after revoming the missing value. However, the proportion of NAs is too large so we have to figure out what is going on here.

```{r}
colSums(is.na(nsch_df)) |>
  sort(decreasing = TRUE) |>
  barplot(cex.names=0.7, las=2)
```
