# Results

## Preprocess

### Load the data

We read `sas7bdat` file downloaded from United States Census Bureau.

"Adult 1 education, household size, and poverty ratio were imputed using sequential regression imputation methods."

```{r}
library(haven)
library(dplyr)
library(tidyr)
nsch_topical <- read_sas(unz('data/nsch_2022e_topical.sas7bdat.zip',
                             'nsch_2022e_topical.sas7bdat'))
nsch_topical <- nsch_topical[nsch_topical$A1_GRADE_IF != 1, ] # imputed rows
```

### Select & Filter

We select to include relevant columns. For details about the meaning of each column, please refer to `data.qmd`. We only study children between age 6-17, so we filter out the `T1` children.

In the end of preprocessing, we stored the data as a csv file called `nsch.csv` inside the data folder

```{r}
parent_ls <- c("A1_ACTIVE", "A1_AGE", "A1_BORN", "A1_EMPLOYED","A1_GRADE",
               "A1_MARITAL", "A1_MENTHEALTH", "A1_PHYSHEALTH", "FAMILY_R",
               "HHCOUNT", "HIGRADE_TVIS", "HOMEEVIC", "HOPEFUL")
experience_ls <- c("ACE1", "ACE10", "ACE11", "ACE12", "ACE3", "ACE4", "ACE5",
                   "ACE6", "ACE7", "ACE8", "ACE9", "BULLIED_R", "BULLY",
                   "EVERHOMELESS", "GRADES")
physical_ls <- c("FASD", "GENETIC", "HOURSLEEP", "PHYSACTIV", "PHYSICALPAIN")
mental_ls <- c("K2Q30A","K2Q31A","K2Q32A", "K2Q33A",
               "K2Q34A","K2Q35A","K2Q37A","K2Q38A")
treatment_ls <- c("K4Q22_R","K4Q23")
nsch_df <- nsch_topical |>
  select(all_of(c("HHID", "FORMTYPE", parent_ls, experience_ls,
                  physical_ls, mental_ls, treatment_ls))) |>
  filter(FORMTYPE != "T1")

write.csv(nsch_df[,-2], "data/nsch.csv", row.names = FALSE)
```

### Missing Value Analysis

```{r}
nsch_df <- read.csv("data/nsch_dropna.csv") # See Chapter 2
```

### Change to Descriptive Column Names

-   K2Q30A ‐ Learning Disability

-   K2Q31A ‐ ADD/ADHD

-   K2Q32A ‐ Depression

-   K2Q33A ‐ Anxiety

-   K2Q34A ‐ Behavior Problems

-   K2Q35A ‐ Autism ASD

-   K2Q37A ‐ Speech Disorder

-   K2Q38A ‐ Tourette Syndrome

```         
my_dataframe <- my_dataframe %>% 
        rename("id" = "c1",
               "pages" = "c2",
               "name" = "c3")`
```

## Guardian's Influence on Children's Mental Health

### Full-time Employment + Education

```{r}
par_grade_employ <- nsch_df |>
  mutate(Fully_Employed = ifelse(nsch_df$A1_EMPLOYED == 1, "Yes", "No")) |>
  mutate(Child_mental_struggle = ifelse(rowSums(nsch_df[, mental_ls] == 1) > 0,
                                      "Yes", "No")) |>
  group_by(Fully_Employed, A1_GRADE, Child_mental_struggle)|>
  summarize(Freq = n())
  
library(vcd)
mosaic(Child_mental_struggle ~ Fully_Employed + A1_GRADE,
       data=par_grade_employ, direction=c("h", "v", "h")) 
```

**A1_EMPLOYED ‐ Adult 1 ‐ Current Employment Status (T1 T2 T3)**

Which of the following best describes your current employment status?

1 = 8th grade or less

2 = 9th‐12th grade; No diploma

3 = High School Graduate or GED Completed

4 = Completed a vocational, trade, or business school program

5 = Some College Credit, but No Degree

6 = Associate Degree (AA, AS)

7 = Bachelor's Degree (BA, BS, AB)

8 = Master's Degree (MA, MS, MSW, MBA)

9 = Doctorate (PhD, EdD) or Professional Degree (MD, DDS, DVM, JD)

### WIthin children with mental struggles, what's proportion of each problem with respect to A1_MENTHEALTH, facet by FAMILY_R

```{r}
par_mental_famstruc_issues <- nsch_df |>
  mutate(Child_mental_struggle = ifelse(rowSums(nsch_df[, mental_ls] == 1) > 0,
                                      "Yes", "No")) |>
  filter(Child_mental_struggle == "Yes") |>
  pivot_longer(cols = K2Q30A:K2Q38A, names_to = "Issues") |>
  select(all_of(c("FAMILY_R", "A1_MENTHEALTH", "Issues", "value"))) |>
  filter(value == 1) |>
  group_by(FAMILY_R, A1_MENTHEALTH, Issues) |>
  summarize(Freq = n())
```

```{r}
library(ggplot2)
plot_df <- par_mental_famstruc_issues |> group_by(FAMILY_R, A1_MENTHEALTH) |> 
  mutate(Total = sum(Freq)) |> ungroup()
ggplot(plot_df, aes(x = A1_MENTHEALTH, y = Issues)) +
  geom_tile(aes(fill = (Freq/Total)), color = "white") +
  coord_fixed() + 
  scale_fill_gradient(low = "white", high = "red") +
  facet_wrap(~FAMILY_R, ncol = 4) + theme_heat
```

**FAMILY_R ‐ Family Structure (T1 T2 T3)**

1 = Two biogical/adoptive parents, currently married

2 = Two biogical/adoptive parents, not currently married

3 = Two parents (at least one not biological/adoptive), currently married

4 = Two parents (at least one not biological/adoptive), not currently married

5 = Single mother

6 = Single father

7 = Grandparent household

8 = Other relation					

**A1_MENTHEALTH ‐ Adult 1 ‐ Mental or Emotional Health (T1 T2 T3)**

In general, how is your mental or emotional health?

1 = Excellent

2 = Very Good

3 = Good

4 = Fair

5 = Poor

### Treatment

```{r}
library(forcats)
nsch_df |>
  mutate(Treatment = ifelse(nsch_df$K4Q22_R == 1, "Yes", "No")) |>
  pivot_longer(mental_ls, names_to = "Issues") |>
  filter(value == 1) |>
  select(all_of(c("Issues", "Treatment"))) |>
  ggplot(aes(x = fct_infreq(Issues), fill = Treatment)) +
  geom_bar() + labs(x = "Issues")
  
```

```{r}
nsch_df |>
  mutate(Treatment = ifelse(nsch_df$K4Q22_R == 1, "Yes", "No")) |>
  pivot_longer(mental_ls, names_to = "Issues") |>
  filter(value == 1) |>
  select(all_of(c("A1_AGE", "Issues", "Treatment"))) |>
  ggplot(aes(x = A1_AGE, y=Treatment)) +
  geom_boxplot()
```
